---
title: "Cleaning_data"
output: html_document
date: "2023-06-08"
---

Libraries

```{r}
library(tidyverse)
library(plyr)
library(dplyr)
library(childsds)
library(MatchIt)

```

Reading data

```{r}
data <- read.csv(r"(C:\Users\User\OneDrive - University of Wyoming\Desktop - Copy\Lab\abcd\0_Working\Clean data\3_Fullest_clean.csv)")
```

#Rename BMI col
Found the col #s

```{r}
#first, find the col #s
names(data)

#6 - age
#7 - household.income 
#9 - bmi
#13 - year
#14 - sex
#15-162 - ROIs


colnames(data[14:162])
colnames(data[6])
colnames(data[7])
colnames(data[9])
colnames(data[13])

#rename bmi
names(data)[9] = "bmi"

```

#Dropping all NAs from cortical regions, age, household.income sex and BMI
##Dropping is based on data from baseline
#bcs no data for household.income for year 2 -> will drop all year 2 data

```{r}
#first create an empty df to write the output df to
dna1 <- data.frame(data)
dna1 <- subset (dna1, dna1$id == "1")

#drop NAs - ONLY HAS BASELINE DATA, NEED TO ADD YEAR 2
for (i in 1:nrow(data)) {
  row <- data[i, ]
  if (row[, 13] == "baseline_year_1_arm_1") {
  dna1 <- rbind(dna1, row[complete.cases(row [, c(6, 7, 9, 11, 13:162)]),])   
  } 
}

#74167 rows > 9083 rows


#-----------------------------------------------------------------------------
#Old dropping NAs method
#dna <- data[complete.cases(data[, c(4, 9, 11, 14:162)]),]

#74167 rows > 15662



```
Setting up 2nd year data
#adding missing income and race_ehnicity data
#rearranging the col#s correctly
#dropping NAs as in baseline
```{r}
data2 <- subset(data, year == "2_year_follow_up_y_arm_1")
#10414 rows
missing_data <- dna1[, c("id", "household.income", "race_ethnicity")]
data2 <- join(data2, missing_data, by = "id")
#10414 rows

#data2 <- merge(data2, missing_data, by = "id")
#merge drops the ids that don't exist in some of these 2 dataframes

#join added the 2 columns just at the end of the dataframe
#now wanna place the columns as in dna1
#7 - household.income
#11 - race_ethnicity
names(data2)
data2 <- data2[, c(-7, -11)] #deleting empty income and race
#162 rows
data2 <- data2[, c(1:6, 161, 7:9, 162, 10:160)]
names(data2)
#162 rows
#Dropping NAs
dna2 <- data2[complete.cases(data2 [, c(6, 9, 14:162)]),]
#4605

```

#Merging baseline and 2nd year data
(bcs 2nd year data has dropped out when I was filtering NAs)

```{r}
dna <- rbind(dna1, dna2)
#13688 rows

```


#Add mean centered age
#deleted income col - it was some household.incom.bl col I renamed to income by mistake
```{r}
dna$mcage <- scale(dna$age, scale = FALSE)

#deleted income col - it was some household.incom.bl col I renamed to income by mistake
dna = subset(dna, select = -income)
```

#Add BMI%
```{r}

#names(dna)
dna$agey <- dna$age/12

dna$bmi_perc <- childsds::sds(dna$bmi,
                    age = dna$agey,
                    sex = dna$sex_at_birth, male = "M", female = "F",
                    ref = cdc.ref,
                    item = "bmi",
                    type = "perc")

dna$bmi_perc <- dna$bmi_perc*100
```

#Filter out underweights
#Add BMI% groups

```{r}
#filter out underweights

dna <- dna %>% filter(bmi_perc > 5)
summary(dna$bmi_perc)

#add new column of bmi categories
dna$bmi_perc_group <- cut(dna$bmi_perc,
                          breaks=c(5, 85, 95, 100),
                          labels=c('Recommended', 'Overweight', 'Obese'))
summary(as.factor(dna$bmi_perc_group))

#Recommended  Overweight       Obese
#2163        2073        8919 
```
#Save full dataset

```{r}
#renamed year back to event_name (to minimize renamings in other scripts)
names(dna)
names(dna)[12] <- "event_name"

#13155 rows
write.csv(dna, r"(C:\Users\User\OneDrive - University of Wyoming\Desktop - Copy\Lab\abcd\0_Working\Clean data\Finalized\Full_bmiperc_wide.csv)", row.names=FALSE)
```


#Now divide into base and follow-up datasets
#Save them

```{r}

summary(as.factor(dna$event_name)) #as factor bcs these variables are categorical

#2_year_follow_up_y_arm_1    baseline_year_1_arm_1 
#                    4436                     8719 

#choosing the baseline data
dna1 <- subset(dna, dna$event_name == "baseline_year_1_arm_1")
summary(as.factor(dna1$event_name))

#baseline_year_1_arm_1 
#                 8719 

write.csv(dna1, r"(C:\Users\User\OneDrive - University of Wyoming\Desktop - Copy\Lab\abcd\0_Working\Clean data\Finalized\Baseline_bmiperc_wide.csv)", row.names=FALSE)

#choosing follow-up data                    
dna2 <- subset(dna, dna$event_name == "2_year_follow_up_y_arm_1")
summary(as.factor(dna2$event_name))

#2_year_follow_up_y_arm_1 
#                    4436


write.csv(dna2, r"(C:\Users\User\OneDrive - University of Wyoming\Desktop - Copy\Lab\abcd\0_Working\Clean data\Finalized\Followup_bmiperc_wide.csv)", row.names=FALSE)

```
#Making a dataset of subs with both baseline and follow-up data
#saving

```{r}
#making a dataset of subs with  both baseline and follow-up data
full_subs <- merge(dna1, dna2, by = "id")
#4340 rows, 329 cols
full_subs$id.y <- full_subs$id
#4340 rows, 330 cols
summary(as.factor(full_subs$event_name.y))

write.csv(full_subs, r"(C:\Users\User\OneDrive - University of Wyoming\Desktop - Copy\Lab\abcd\0_Working\Clean data\Finalized\Subs_with_both_years_bmiperc_wide.csv)", row.names=FALSE)

```




#Matching datasets between groups
Year1 
```{r}
summary(as.factor(dna1$bmi_perc_group))
#Obese  Overweight Recommended 
#       1424        1332        5963 


#I didn't figure out how to make three equal samples
#So I will have to use the matchit function
#It requires 2 samples 
#Thus I am dropping the overweight
mdna1 <-dna1
mdna1 <-subset(dna1, dna1$bmi_perc_group != "Overweight")

#still has Overweight group which makes our group non-binary
mdna1$bmi_perc_group <- droplevels(as.factor(mdna1$bmi_perc_group))
summary(as.factor(mdna1$bmi_perc_group))  
#Recommended       Obese 
#       5963         1424 
#matching the bmi groups
mdna1 <- matchit(as.factor(bmi_perc_group) ~ age + sex_at_birth, data = mdna1)
summary(mdna1)
mdna1 <- match.data(mdna1)
summary(mdna1$bmi_perc_group)
#Recommended       Obese 
#        1424         1424 
```

Year 2
```{r}
#I didn't figure out how to make three equal samples
#So I will have to use the matchit function
#It requires 2 samples 
#Thus I am dropping the overweight
mdna2 <- dna2
mdna2 <- subset(dna2, dna2$bmi_perc_group != "Overweight")

#still has Ovwerweight group which makes our group non-binary
mdna2$bmi_perc_group <- droplevels(as.factor(mdna2$bmi_perc_group))
summary(as.factor(mdna2$bmi_perc_group))  
#Recommended       Obese 
#       739         2956 

#matching the bmi groups
mdna2 <- matchit(as.factor(bmi_perc_group) ~ age + sex_at_birth, data = mdna2)
mdna2 <- match.data(mdna2)
summary(mdna2$bmi_perc_group)
#Recommended       Obese 
#        739         739 

```
Saving matched datasets

```{r}
write.csv(mdna1, r"(C:\Users\User\OneDrive - University of Wyoming\Desktop - Copy\Lab\abcd\0_Working\Clean data\Finalized\MATCHED_Baseline_bmiperc_wide.csv)", row.names=FALSE)
write.csv(mdna2, r"(C:\Users\User\OneDrive - University of Wyoming\Desktop - Copy\Lab\abcd\0_Working\Clean data\Finalized\MATCHED_Followup_bmiperc_wide.csv)", row.names=FALSE)

```




#READ TO AVOID RERUNNING THE ABOVE LOOP

```{r}
#dna <- read.csv(r"(C:\Users\User\OneDrive - University of Wyoming\Desktop - Copy\Lab\abcd\0_Working\Clean data\Finalized\Full_bmiperc_wide.csv)")

dna1 <- read.csv(r"(C:\Users\User\OneDrive - University of Wyoming\Desktop - Copy\Lab\abcd\0_Working\Clean data\Finalized\Baseline_bmiperc_wide.csv)")
dna2 <- read.csv(r"(C:\Users\User\OneDrive - University of Wyoming\Desktop - Copy\Lab\abcd\0_Working\Clean data\Finalized\Followup_bmiperc_wide.csv)")

#mdna1 <- read.csv(r"(C:\Users\User\OneDrive - University of Wyoming\Desktop - Copy\Lab\abcd\0_Working\Clean data\Finalized\MATCHED_Baseline_bmiperc_wide.csv)")
#mdna2 <- read.csv(r"(C:\Users\User\OneDrive - University of Wyoming\Desktop - Copy\Lab\abcd\0_Working\Clean data\Finalized\MATCHED_Followup_bmiperc_wide.csv)")



full_subs <- read.csv(r"(C:\Users\User\OneDrive - University of Wyoming\Desktop - Copy\Lab\abcd\0_Working\Clean data\Finalized\Subs_with_both_years_bmiperc_wide.csv)")

```

#Dropping overweights (for matching between the groups)

```{r}

dna1_nov <- subset(dna1, dna1$bmi_perc_group != "Overweight")
dna2_nov <- subset(dna2, dna2$bmi_perc_group != "Overweight")

#still has Ovwerweight group which makes our group non-binary
dna1_nov$bmi_perc_group <- droplevels(as.factor(dna1_nov$bmi_perc_group))
summary(as.factor(dna1_nov$bmi_perc_group))  
# Obese     Recommended 
#       1424        5963  
       
dna2_nov$bmi_perc_group <- droplevels(as.factor(dna2_nov$bmi_perc_group))
summary(as.factor(dna2_nov$bmi_perc_group))  

#Obese      Recommended 
#        739        2956   
```


##MAKING LONG DATA FOR REP MANOVA

#First, get the obese-rec groups
#then add 2nd year data

```{r}
IDs <- merge(dna1_nov, dna2_nov, by="id")

IDs <- as.data.frame(IDs[, -2:-329])
names(IDs[, 1])
#turned out the column didn't have a name
#so I'm naming it bcs that's needed to merge 2 dfs by col name
colnames(IDs)[1] <-  "id"

#3270 rows
```


Match the ob-rec groups in year 1
then will add the 2nd year data
```{r}
dna1_nov <- merge(dna1_nov, IDs, by = "id")

#3270 rows
#matching the bmi groups
mdna1_nov <- matchit(as.factor(bmi_perc_group) ~ age + sex_at_birth, data = dna1_nov)
mdna1_nov <- match.data(mdna1_nov)
summary(mdna1_nov$bmi_perc_group)
#Recommended       Obese 
#     570        570  
#1140 total, 168 cols

#drop added cols
mdna1_nov <- mdna1_nov[, c(-166:-168)] 
#1140 total, 165 cols
```

Now need IDs of these subs to add 2nd year data


```{r}
mIDs <- merge(mdna1_nov, dna2_nov, by="id")

mIDs <- as.data.frame(mIDs[, -2:-329])
names(mIDs[, 1])
#turned out the column didn't have a name
#so I'm naming it bcs that's needed to merge 2 dfs by col name
colnames(mIDs)[1] <-  "id"

#3270 rows
```

Merging the base and FU data in long table, not wide 

```{r}
#rois1 <- merge(dna1, IDs)
#rois2 <- merge(dna2, IDs)
#then binded them
#complied into 1 line to avoid too many vars 

mfull <-rbind(merge(mdna1_nov, mIDs), merge(dna2_nov, mIDs))
summary(as.factor(mfull$event_name))
#2_year_follow_up_y_arm_1    baseline_year_1_arm_1 
#                    1140                     1140 
summary(as.factor(mfull$bmi_perc_group))
# Obese Recommended 
#      1109        1171
#2280 rows, 165 cols  

```



#Making long data for rep Manova

```{r}
names(mfull)

#deleted unnecessary
mfull <- mfull[, c(-2:-4, -8:-9, -163:-164)]
#165 cols > 158
names(mfull)

#re-ordered to keep the table cleaner
mfull <- mfull[, c(1:8, 157:158, 9:156)]

#2788, 158

total_long <- gather(mfull, ROI, BOLD, g.and.s.frontomargin.lh:s.temporal.transverse.rh, factor_key=TRUE)

total_long <- total_long[order(total_long$id, total_long$ROI, total_long$event_name),]
summary(as.factor(total_long$event_name))

```

#Coding factors as factors

```{r}
names(total_long)
lapply(total_long, class)
total_long[, c(2, 4:8, 10:11)] <- lapply(total_long[, c(2, 4:8, 10:11)],as.factor)

lapply(total_long, class)
```



Saving the matched long data for subs with 2 years
```{r}
write.csv(total_long, r"(C:\Users\User\OneDrive - University of Wyoming\Desktop - Copy\Lab\abcd\0_Working\Clean data\Finalized\MATCHED_subs_with_2yeardata_bmiperc_long.csv)", row.names=FALSE)
```



