---
title: "ROIS_dividing"
output: html_document
date: "2023-06-12"
---
Libraries

```{r}
library(plyr)
library(dplyr)
library(tidyverse)
library(MatchIt)
```


Reading data

```{r}
dna1 <- read.csv(r"(C:\Users\User\OneDrive - University of Wyoming\Desktop - Copy\Lab\abcd\0_Working\Clean data\Finalized\Baseline_bmiperc_wide.csv)")
dna2 <- read.csv(r"(C:\Users\User\OneDrive - University of Wyoming\Desktop - Copy\Lab\abcd\0_Working\Clean data\Finalized\Followup_bmiperc_wide.csv)")
```

#Dropping overweights (for matching between the groups)

```{r}

dna1_nov <- subset(dna1, dna1$bmi_perc_group != "Overweight")
dna2_nov <- subset(dna2, dna2$bmi_perc_group != "Overweight")

#still has Ovwerweight group which makes our group non-binary
dna1_nov$bmi_perc_group <- droplevels(as.factor(dna1_nov$bmi_perc_group))
summary(as.factor(dna1_nov$bmi_perc_group))  
# Obese     Recommended 
#       1424        5963  
       
dna2_nov$bmi_perc_group <- droplevels(as.factor(dna2_nov$bmi_perc_group))
summary(as.factor(dna2_nov$bmi_perc_group))  

#Obese      Recommended 
#        739        2956   
```


#First, get the obese-rec groups
#then add 2nd year data

```{r}
IDs <- merge(dna1_nov, dna2_nov, by="id")

IDs <- as.data.frame(IDs[, -2:-329])
names(IDs[, 1])
#turned out the column didn't have a name
#so I'm naming it bcs that's needed to merge 2 dfs by col name
colnames(IDs)[1] <-  "id"

#3270 rows
```



Match the ob-rec groups in year 1
then will add the 2nd year data
```{r}
dna1_nov <- merge(dna1_nov, IDs, by = "id")

#3270 rows
#matching the bmi groups
mdna1_nov <- matchit(as.factor(bmi_perc_group) ~ age + sex_at_birth, data = dna1_nov)
mdna1_nov <- match.data(mdna1_nov)
summary(mdna1_nov$bmi_perc_group)
#Recommended       Obese 
#     570        570  
#1140 total, 168 cols

#drop added cols
mdna1_nov <- mdna1_nov[, c(-166:-168)] 
#1140 total, 165 cols
```
Arranging the cols
#Selecting ROIS
```{r}
#names(mfull)

#deleted unnecessary
mdna1_nov <- mdna1_nov[, c(-2:-4, -8:-9, -163:-164)]
#165 cols > 158
names(mdna1_nov)

#re-ordered to keep the table cleaner
mdna1_nov <- mdna1_nov[, c(1:8, 157:158,9:156)]

#2788, 158


#deleted unnecessary
dna2_nov <- dna2_nov[, c(-2:-4, -8:-9, -163:-164)]
#165 cols > 158
names(dna2_nov)

#re-ordered to keep the table cleaner
dna2_nov <- dna2_nov[, c(1:8, 157:158,9:156)]

#2788, 158

#-------------------------------------------------------------------------------

names(mdna1_nov)

mdna1_nov <- mdna1_nov[, c(-166:-168)]

mdna1_nov <- mdna1_nov[, c(-15:-165)]
dna2_nov <- dna2_nov [, c(-15:-165)]


```



Now need IDs of these subs to add 2nd year data


```{r}
mIDs <- merge(mdna1_nov, dna2_nov, by="id")

mIDs <- as.data.frame(mIDs[, -2:-329])
names(mIDs[, 1])
#turned out the column didn't have a name
#so I'm naming it bcs that's needed to merge 2 dfs by col name
colnames(mIDs)[1] <-  "id"

#3270 rows
```
Merging the base and FU data in long table, not wide 

```{r}
#rois1 <- merge(dna1, IDs)
#rois2 <- merge(dna2, IDs)
#then binded them
#complied into 1 line to avoid too many vars 

mfull <-rbind(merge(mdna1_nov, mIDs), merge(dna2_nov, mIDs))
summary(as.factor(mfull$event_name))
#2_year_follow_up_y_arm_1    baseline_year_1_arm_1 
#                    1140                     1140 
summary(as.factor(mfull$bmi_perc_group))
# Obese Recommended 
#      1109        1171
#2280 rows, 165 cols  

```



#Making long data for rep Manova

```{r}


total_long <- gather(mfull, ROI, BOLD, g.and.s.frontomargin.lh:g.and.s.subcentral.lh, factor_key=TRUE)

total_long <- total_long[order(total_long$id, total_long$ROI, total_long$event_name),]
summary(as.factor(total_long$event_name))

```

#Coding factors as factors

```{r}
names(total_long)
lapply(total_long, class)
total_long[, c(2, 4:8, 10:11)] <- lapply(total_long[, c(2, 4:8, 10:11)],as.factor)

lapply(total_long, class)
```


Saving the matched long data for subs with 2 years
```{r}
write.csv(total_long, r"(C:\Users\User\OneDrive - University of Wyoming\Desktop - Copy\Lab\abcd\0_Working\Clean data\Finalized\ROIS for RepMANOVA\1.csv)", row.names=FALSE)
```

